---
title: "Functions in R from Advanced R"
output:
  html_document: default
  html_notebook: default
---

The three components of a function

```{r}
f <- function(x) x^2
f
formals(f) ## the arguments to the function
body(f)    ## the "stuff" in middle, usually between {}
environment(f) ## where  your function going to look for value
               ## of objects

```
These assignments can be used to modify these attributes of a function.  Functions have other attributes as well (like many other objects, remember ! functions are objects)

One attribute is srcref which is short for source reference *(look up later)*



**Privative Functions**
Primative functions do not have a body attribute because they call C+ routines directly. Asking for body of a primative function will return NULL as will requests for formals and environment.
```{r}
x<-1:10
sum(x)
body(sum)
```
Primative functions are only found in the BASE R package.  They do not behave entirely the same as the other functions in R. *(research primative functions argument matching)* but are more efficient.

**Exercises**
The function str() or is.function() will tell you if something is a funtion

```{r}
str(f)
is.function(f)
str(sum)
is.function(sum)
```

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
funs[1:5]

is.function(funs[[2]])
length(formals(funs[[2]]))
args<-sapply(funs,function(x) length(formals(x)))## x a function here?
args[1:10]
(args2<-max(args))

args[args==22]### scan
funs[args==0]

funs[1]
names(funs[1])
str(funs[1])
funs[["-"]]


```


Which base function has the most arguments? *scan has 22 arguments*

How many base functions have no arguments? What’s special about those functions? *Many of them are Primative functions but not all are.*

How could you adapt the code to find all primitive function
*not sure*

---











